\bf Title here

### 2020-09-17 paper/paper.Rmd
some coarse writing on pca & animated & tour

library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)
library(tourr)
```

@@

In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are of dimension $p \times d$ with $p$ usually greater than two and hence can't be visualised in a 2D plot. An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA). The visualisation can thus be defined as

@@

Figure 5 shows the searching space plot of a geodesic search, colored by different searching stages. Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions. The best one is picked and an exact search finds the stepsize that maximises the index value and outputs the corresponding projection basis.

```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
                holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
 bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
  bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
  mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
  compute_pca() %>%
  `$`(aug)

dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
#dt %>% explore_proj_pca(facet = algo, axis = 2)
```

@@

all the plots above can be animated if wanted - importance is how the data is moving around

show an example of some snap shots  of the animation and provide a link... for animated pca plot

```{r animated-pca, fig.height = 8, fig.cap = "\\label{animated-pca} 16 frames from animated PCA plot ." }
rl <- lapply(list.files(here::here("figures/animated_pca/"),full.names = TRUE)[1:16], png::readPNG)
gl <-  lapply(rl, grid::rasterGrob)
gridExtra::grid.arrange(grobs=gl)
```

@@

show the path of the tour on a sphere. (full data dimension animation...)

include the tour animation (geozoo)

>>>>

library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)
library(tourr)
library(gganimate)
```

@@

The projection bases $\mathbf{A}_{p \times d}$ live on the surface of a $p \times d$ dimension sphere, where this dimension can easily go from 5 to 10. This makes viewing the original space hard. An easier alternative is to view the reduced space via some dimensional reduction methods i.e. Principal component analysis. The visualisation can thus be defined as

@@

Figure \ref{toy-pca} shows the first two principal components of two searching paths. A Hollow sphere with the same dimension is overlaid to emphasize the imposition of orthonormal constraint. Starting from the same position, two searching methods takes different paths to get to the two final bases with the same data projection $Y$ (the two correspond to a rotation of the data in the projection). The distinction between the two is that `search_better()` also involve some random evaluation in the sphere due to its sampling process while `search_geodesic()` doesn't. This stochastic rather than deterministic approach can be useful in more complex scenario.

```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
# dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
#                 holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
#  bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
#   bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
#   mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
#   compute_pca() %>%
#   `$`(aug)
# 
# 
# dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
# #dt %>% explore_proj_pca(facet = algo, axis = 2)

path1 <- holes_1d_geo$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
path2 <- holes_1d_better$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
set.seed(1)
sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
theoretical <- matrix(c(0, 1, 0, 0, 0,
                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)

dt <- rbind(sphere, path1, path2, theoretical)
pca <- dt %>% prcomp()
#summary(pca)
pca_pred <- predict(pca) %>%
  as_tibble() %>%
  mutate(method = c(rep("sphere", nrow(sphere)),
                    holes_1d_geo$method,
                    holes_1d_better$method,
                    rep("theoretical", 2))) %>%
  group_by(method) %>%
  mutate(id = row_number(),
         id = ifelse(method == "sphere", 1, id)) %>%
  ungroup()

p <- pca_pred %>%
  ggplot(aes(x = PC1, y = PC2, color = method))+
  geom_point() +
  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
             size = 10) + 
  theme(aspect.ratio = 1) +
  scale_color_botanical(palette = "cherry")

p
```

@@

An animated plot can be useful because it allows us to track the progression of the optimsiation, especially in the case of inspecting the search space.

<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->

<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->

<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->

<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->

<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->

```{r cache = TRUE, eval=FALSE}
# rl <- lapply(list.files(here::here("figures/animated_pca_1d/"),full.names = TRUE)[62:78], png::readPNG)
# gl <-  lapply(rl, grid::rasterGrob)
# gridExtra::grid.arrange(grobs=gl)
p +
  transition_states(id) +
  shadow_mark()
```

@@

As mentioned previously, the orthonormal constraint regulates the projection basis to be living on the surface of a unit sphere. The visualisation of the projection basis on its original $p \times d$ dimension can be done through an animated rotation of the $p \times d$ diemnsion sphere with the searching path overlaid. The high dimensional geometry can be simulated using the package `geozoo` and the animated rotation can be done via a regular grand tour or a slice tour [@laa2020slice].

Compared with the PCA plot, this animated rotation display requires skills of the viewer to preceive the high dimensional space in rotation and what the viewers gain is a much holistic view of the path in the space in the sense that PCA does a further projection on the projection basis along with the hollow sphere, which gives the view that maixmises the variance. While the animated rotation display gives the view from different anlges and the PCA plot consitutes one frame from the animated rotaiton display.

### 2020-09-17 paper/paper.tex
some coarse writing on pca & animated & tour

In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
are of dimension \(p \times d\) with \(p\) usually greater than two and
hence can't be visualised in a 2D plot. An option to explore the
searching space of these bases is to explore a reduced space via
principal component analysis (PCA). The visualisation can thus be
defined as

@@

Figure 5 shows the searching space plot of a geodesic search, colored by
different searching stages. Starting from the start basis, a directional
search is conducted in a narrow neighbourhood on five random directions.
The best one is picked and an exact search finds the stepsize that
maximises the index value and outputs the corresponding projection
basis.

@@

all the plots above can be animated if wanted - importance is how the
data is moving around

show an example of some snap shots of the animation and provide a
link\ldots{} for animated pca plot

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/animated-pca-1.pdf}
\caption{\label{animated-pca} 16 frames from animated PCA plot .}
\end{figure}

@@

show the path of the tour on a sphere. (full data dimension
animation\ldots{})

include the tour animation (geozoo)

@@

\begin{verbatim}
## # A tibble: 10 x 8
##       id basis             index_val info         tries  loop method       alpha
##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 <NA>           0.5
##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     1 search_bett~  NA  
##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA  
## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     3 search_bett~  NA
\end{verbatim}

>>>>

The projection bases \(\mathbf{A}_{p \times d}\) live on the surface of
a \(p \times d\) dimension sphere, where this dimension can easily go
from 5 to 10. This makes viewing the original space hard. An easier
alternative is to view the reduced space via some dimensional reduction
methods i.e.~Principal component analysis. The visualisation can thus be
defined as

@@

Figure \ref{toy-pca} shows the first two principal components of two
searching paths. A Hollow sphere with the same dimension is overlaid to
emphasize the imposition of orthonormal constraint. Starting from the
same position, two searching methods takes different paths to get to the
two final bases with the same data projection \(Y\) (the two correspond
to a rotation of the data in the projection). The distinction between
the two is that \texttt{search\_better()} also involve some random
evaluation in the sphere due to its sampling process while
\texttt{search\_geodesic()} doesn't. This stochastic rather than
deterministic approach can be useful in more complex scenario.

@@

An animated plot can be useful because it allows us to track the
progression of the optimsiation, especially in the case of inspecting
the search space.

@@

As mentioned previously, the orthonormal constraint regulates the
projection basis to be living on the surface of a unit sphere. The
visualisation of the projection basis on its original \(p \times d\)
dimension can be done through an animated rotation of the \(p \times d\)
diemnsion sphere with the searching path overlaid. The high dimensional
geometry can be simulated using the package \texttt{geozoo} and the
animated rotation can be done via a regular grand tour or a slice tour
\citep{laa2020slice}.

Compared with the PCA plot, this animated rotation display requires
skills of the viewer to preceive the high dimensional space in rotation
and what the viewers gain is a much holistic view of the path in the
space in the sense that PCA does a further projection on the projection
basis along with the hollow sphere, which gives the view that maixmises
the variance. While the animated rotation display gives the view from
different anlges and the PCA plot consitutes one frame from the animated
rotaiton display.

@@

\begin{verbatim}
## # A tibble: 10 x 8
##       id basis             index_val info         tries  loop method       alpha
##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 search_bett~   0.5
##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     1 search_bett~  NA  
##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA  
## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     3 search_bett~  NA
\end{verbatim}
